---
title: "Customer Churn Prediction"
author: "Jose Alberto"
date: "08/06/2021"
output: github_document
---

```{r}
rm(list=ls(all=TRUE))
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


Customer Churn

A churn is when a customer stops paying or subscribing to your products or  services.
In other word, is the loss these customers.

Customer churn is one of the most important business metrics.

That’s because the cost of retaining an existing customer is significantly less than the cost of acquiring a new one.

The latter is often referred to as Customer Acquisition Cost (CAC).

Companies use it as a metric to track if a company has a viable business model that can keep profits generating while maintaining a low CAC.

Load R Packages: 

```{r}
library(rmarkdown)
library(htmltools)
library(tidyverse)
library(dplyr)#dplyr is a grammar of data manipulation
library(lattice)
library(survival)
library(Formula)
library(missForest)
library(MLmetrics)
library(kernlab)
library(gridExtra)
library(caTools)#For Logistic regression
library(Amelia)#Amelia II is an R package for the multiple imputation of multivariate incomplete data
library(Hmisc)#imputing missing values
library(dlookr)#Tools for Data Diagnosis, Exploration, Transformation
library(rcompanion)#Functions to Support Extension Education Program Evaluation
library(PerformanceAnalytics)#Econometric Tools for Performance and Risk Analysis
library(corrplot)
library(lubridate)#Make Dealing with Dates a Little Easier
library(fpp)#Data for Forecasting
library(caret)#Classification and Regression Training        
library(caretEnsemble)#Ensembles of Caret Models
library(e1071)#Misc Functions of the Department of Statistics
library(plyr)#Tools for Splitting, Aplitting and Combining Data
library(randomForest)#Random Forest for Classification 
library(ROCR)
```


```{r pressure, echo=FALSE}
training_dataset <- read.csv("datasets/ml_case_training_data.csv", encoding='UTF-8', header = TRUE, na.strings = "")
```

```{r pressure, echo=FALSE}
history_dataset <- read.csv("datasets/ml_case_training_hist_data.csv", encoding='UTF-8', header = TRUE, na.strings = "")
```

```{r pressure, echo=FALSE}
churn_dataset <- read.csv("datasets/ml_case_training_output.csv", encoding='UTF-8', header=TRUE, na.strings = "")
```


```{r}
View(training_dataset)
```



```{r}
training_dataset2 <- training_dataset
```


```{r}
history_dataset2 <- history_dataset
```


```{r}
dim(training_dataset2)
```

```{r}
dim(history_dataset)
```

```{r}
dim(churn_dataset)
```

```{r}
str(training_dataset2)
```


```{r}
str(history_dataset2)
```


```{r}
str(churn_dataset)
```


## Using sapply on columns with missing values:
```{r}
sapply(training_dataset2, function(x) sum(is.na(x)))
```

```{r}
sapply(history_dataset2, function(x) sum(is.na(x)))
```

Compute the total missing values in each columns:

```{r}
colSums(is.na(churn_dataset))
```



### Plot a missinggness map showing where occurs in the dataset:

```{r}
missmap(training_dataset2, 
        legend=TRUE,
        col=c('#ff5c33', '#3399ff'),
        y.cex = 0.5,
        x.cex = 0.5
)
```


That finds duplicate values in a dataset:

```{r}
anyDuplicated(training_dataset)
```

```{r}
anyDuplicated(history_dataset)
```


```{r}
anyDuplicated(churn_dataset)
```


```{r}
summary(training_dataset[,c(5,6,7,13,14,15,16,17,18,19,20,21,22,23)])
```


```{r}
hist(na.pass(training_dataset$forecast_base_bill_ele), 
     col = 'blue',
     xlab = 'forecast_base_bill_ele',
     main = 'Distribution Variable Values of the Forecast Base Bill Ele')
```


```{r}
layout(matrix(c(1,2), nrow=1, ncol=2))
boxplot(training_dataset[training_dataset$forecast_bill_12m>=1,15], main = 'Forecast Bill 12m',pch=16, col = '#99ccff', border = '#00264d')
boxplot(training_dataset[training_dataset$forecast_price_energy_p1!=0,16], main = 'Forecast Price Energy p1',pch=16, col = '#99ffcc', border = '#00264d')
```


```{r}
layout(matrix(c(1,2), nrow=1, ncol=2))
plot_outlier(training_dataset[5])
```

```{r}
layout(matrix(c(1,2), nrow=2, ncol=2))
plot_normality(training_dataset[13])
```



##### Data Cleaning

Impute missing values to 0 in columns activity_new,campaign_disc_ele, channel_sales :


```{r}
training_dataset2[is.na(training_dataset2$activity_new),]$activity_new <- 0
```



```{r}
training_dataset2[is.na(training_dataset2$campaign_disc_ele),]$campaign_disc_ele <- 0
```


```{r}
training_dataset2[is.na(training_dataset2$channel_sales),]$channel_sales <- 0
```


```{r}
training_dataset2[is.na(training_dataset2$origin_up),]$origin_up <- 0
```



 Using impute function from Hmisc package replace with median the columns:
 
```{r}
training_dataset2$forecast_base_bill_ele <- Hmisc::impute(training_dataset2$forecast_base_bill_ele, median)
```


```{r}
training_dataset2$forecast_base_bill_year <- Hmisc::impute(training_dataset2$forecast_base_bill_year, median)
```


```{r}
training_dataset2$forecast_bill_12m <- Hmisc::impute(training_dataset2$forecast_bill_12m, median)
```


```{r}
training_dataset2$forecast_cons <- Hmisc::impute(training_dataset2$forecast_cons, median)
```


```{r}
training_dataset2$forecast_discount_energy  <- Hmisc::impute(training_dataset2$forecast_discount_energy, median)
```


```{r}
training_dataset2$forecast_price_pow_p1  <- Hmisc::impute(training_dataset2$forecast_price_pow_p1, median)
```


```{r}
training_dataset2$forecast_price_energy_p1  <- Hmisc::impute(training_dataset2$forecast_price_energy_p1, median)
```


```{r}
training_dataset2$forecast_price_energy_p2  <- Hmisc::impute(training_dataset2$forecast_price_energy_p2, median)
```


```{r}
training_dataset2$margin_gross_pow_ele <- Hmisc::impute(training_dataset2$margin_gross_pow_ele, median)
```


```{r}
training_dataset2$margin_net_pow_ele <- Hmisc::impute(training_dataset2$margin_net_pow_ele, median)
```


```{r}
training_dataset2$net_margin <- Hmisc::impute(training_dataset2$net_margin, median)
```


```{r}
training_dataset2$pow_max <- Hmisc::impute(training_dataset2$pow_max, median)
```



```{r}
history_dataset2$price_p1_var <-  Hmisc::impute(history_dataset2$price_p1_var, median)
```



```{r}
history_dataset2$price_p2_var <-  Hmisc::impute(history_dataset2$price_p2_var, median)
```


```{r}
history_dataset2$price_p3_var <-  Hmisc::impute(history_dataset2$price_p3_var, median)
```


```{r}
history_dataset2$price_p1_fix <-  Hmisc::impute(history_dataset2$price_p1_fix, median)
```


```{r}
history_dataset2$price_p2_fix <-  Hmisc::impute(history_dataset2$price_p2_fix, median)
```


```{r}
history_dataset2$price_p3_fix <-  Hmisc::impute(history_dataset2$price_p3_fix, median)
```



###### The three forms of correlation presented here are Pearson, Kendall, and Spearman. 

Kendall and Spearman correlation use nonparametric tests. 

The test determining the p-value for Pearson correlation is a parametric test that assumes that data are bivariate normal.


```{r}
correlations_training_dataset2 <- cor(training_dataset2[,c(5,6,7,13,14,15,16,17,18,19,20,21)], method = "spearman")
```


```{r}
corrplot(correlations_training_dataset2, method="circle", type = "full")
```


```{r}
training_dataset2$has_gas <- mapvalues(training_dataset2$has_gas, from = c("f", "t"), to = c("FALSE", "TRUE"))
```


```{r}
training_dataset2$has_gas <- as.logical(training_dataset2$has_gas)
```


Coverting Strings to Date:
```{r}
training_dataset2$date_activ <- as.Date(training_dataset2$date_activ)
```


```{r}
training_dataset2$date_end <- as.Date(training_dataset2$date_end)
```


```{r}
training_dataset2$date_first_activ <- as.Date(training_dataset2$date_first_activ, format="%Y-%m-%d")
```


```{r}
training_dataset2$date_modif_prod <- as.Date(training_dataset2$date_modif_prod)
```


```{r}
training_dataset2$date_renewal <- as.Date(training_dataset2$date_renewal)
```


```{r}
history_dataset2$price_date <- as.Date(history_dataset2$price_date)
```


Forecasting Times Series :

```{r}
ts_forecast_base_bill_year <- ts(unique(training_dataset2$forecast_base_bill_year), start=c(1970,1), end = c(2021,12),frequency=12)
```


```{r}
ts.plot(ts_forecast_base_bill_year, 
     col="#004080", 
     xlab="Year",
     ylab="Electricity Bill Baseline")
```

